---
title: "groundwaterconductivity"
author: "Patrick van Beelen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{groundwaterconductivity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(groundwaterconductivity)
```

## Introduction

Whereas the README.Rds file gives you information about what you can do with the groundwaterconductivity package. This manual will show you how to do it. The method is described in a Dutch report by Stuyfzand, P. (1987). Een zeer nauwkeurige berekening van het elektrisch geleidingsvermogen, ter controle en aanvulling van wateranalyses: 2e versie. Rijswijk, KIWA. Table 3.1 of this report is used to test the calculations. The table 3.1 is shown in data/Table.csv

## Data input

The groundwater conductivity calculations require proper names and dimensions of the input parameters needed to calculate the conductivity. Table 1 below presents the name, content and dimension of the parameters needed. The first rows from rownumber to Calculated conductivity are also present in Table 3.1 of the Stuyfzand report and are used for the quality test of the calculations. Stuyfzand did not use aluminum, iron, phosphate, manganese and zinc to validate his method. The Dutch Minerals Monitoring Policy Program is using the groundwater conductivity calculations presented here for decades. This program is abbreviated to LMM in Dutch and has its own parameter names and dimensions. In most cases milligrams/liter is used except for a few differences mentioned in column LMM dimension.

Table 1 measured parameters (`Table`):

+------------+-------------------------+------------+------------+------------+
| name       | content                 | dimension  | LMM        | LMM        |
|            |                         |            |            |            |
|            |                         |            | name       | dimension  |
+:===========+:========================+:===========+============+============+
| no         | rownumber               |            | myrownames |            |
+------------+-------------------------+------------+------------+------------+
| cl         | Choride                 | mg/L       | xcl        |            |
+------------+-------------------------+------------+------------+------------+
| hco3       | Bicarbonate             | mg/L       | xhco3      |            |
+------------+-------------------------+------------+------------+------------+
| so4        | Sulfate                 | mg/L       | xso4       |            |
+------------+-------------------------+------------+------------+------------+
| no3        | Nitrate                 | mg/L       | xno3       |            |
+------------+-------------------------+------------+------------+------------+
| co3        | Carbonate               | mg/L       | xco3       |            |
+------------+-------------------------+------------+------------+------------+
| h          | pH                      |            | xhv        |            |
+------------+-------------------------+------------+------------+------------+
| na         | Sodium                  | mg/L       | xna        |            |
+------------+-------------------------+------------+------------+------------+
| k          | Potassium               | mg/L       | xk         |            |
+------------+-------------------------+------------+------------+------------+
| ca         | Calcium                 | mg/L       | ca         |            |
+------------+-------------------------+------------+------------+------------+
| mg         | Magnesium               | mg/L       | xmg        |            |
+------------+-------------------------+------------+------------+------------+
| nh4        | Ammonium                | mg/L       | xnh4       |            |
+------------+-------------------------+------------+------------+------------+
| k20        | Measured conductivity   | µS/cm      | xecv       | mS/m       |
+------------+-------------------------+------------+------------+------------+
| rk20       | Calculated conductivity | µS/cm      | ec25       | mS/m       |
+------------+-------------------------+------------+------------+------------+
| al         | Aluminum                | mg/l       | xal        | µg/L       |
+------------+-------------------------+------------+------------+------------+
| fe         | Iron                    | mg/L       | xfe        |            |
+------------+-------------------------+------------+------------+------------+
| po4        | Phosphate               | mg/L       | xpo4       | mg P/L     |
+------------+-------------------------+------------+------------+------------+
| mn         | Manganese               | mg/L       | xmn        |            |
+------------+-------------------------+------------+------------+------------+
| zn         | Zinc                    | mg/L       | xzn        | µg/L       |
+------------+-------------------------+------------+------------+------------+

Stuyfzand measured the conductivity at 20°C and expressed it in micro Siemens/centimeter. At the LMM the conductivity is determined at 25°C and is expressed in milliSiemens/meter. In fact it is measured in the field and a temperature correction is made to get the measured conductivity at 25°C. The conductivity increases at higher temperatures and Stuyfzand calculates the k20 from the conductivity kt at a given temperature t °C using: k20 = kt \* \[1+0.023 \* (20-t)\]

Hence, the calculated conductivity rk20 at 20°C can be converted to the calculated conductivity ec25 at 25°C using: ec25 = 0.1 \* rk20/ \[1-0.023 \* (25-20)\]

The data input results in a standard broad LMM style input file like data/Table_Stuyfzand_LMM_broad_input_dataframe.rds It has all the necessary input parameters mentioned in table 1. The parameters that are not provided are set to zero except for the pH which is set to 7. This makes it possible to check whether the input parameters are correct.

## Data output

The data output results in a standard broad LMM style output file like data/Table_Stuyfzand_LMM_broad_output_dataframe.rds It has 73 parameters mentioned in table 2.

Table 2 Output headings

+----------------------+------------+---------------------------------------------------------------------------------+
| name                 | dimension  | content                                                                         |
+======================+============+=================================================================================+
| myrownames           |            |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| no                   |            |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xal                  | µg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xca                  | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xcl                  | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xfe                  | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xhv                  | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xk                   | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xmg                  | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xmn                  | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xna                  | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xnh4                 | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xno3                 | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xpo4                 | mg P/L     |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xso4                 | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xzn                  | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xhco3                | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xco3                 | mg/L       |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| k20meas              | µS/cm      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| k20                  | µS/cm      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| orderK20             |            |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| hco3                 | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| cl                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| so4                  | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| no3                  | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| co3                  | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| h3o                  | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| na                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| k                    | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| ca                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| mg                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| nh4                  | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| fe                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| mn                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| al                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| zn                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| po4                  | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| pos                  | meq/L      | positive ions                                                                   |
+----------------------+------------+---------------------------------------------------------------------------------+
| neg                  | meq/L      | negative ions                                                                   |
+----------------------+------------+---------------------------------------------------------------------------------+
| ib                   |            | ion balance                                                                     |
+----------------------+------------+---------------------------------------------------------------------------------+
| oh                   | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| mu                   |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| sqmu                 |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| gam2                 |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| san                  | meq/L      | sum anions                                                                      |
+----------------------+------------+---------------------------------------------------------------------------------+
| skat                 | meq/L      | sum kations                                                                     |
+----------------------+------------+---------------------------------------------------------------------------------+
| alM                  |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| so4M                 |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| DO                   |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| tmp                  |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| also4                |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| alF                  |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| aloh                 |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| aloh2                |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| alZ                  |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| san2                 | meq/L      | sum anions corrected                                                            |
+----------------------+------------+---------------------------------------------------------------------------------+
| skat2                | meq/L      | sum cations corrected                                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| sgem                 | meq/L      | (san2+skat2)/2                                                                  |
+----------------------+------------+---------------------------------------------------------------------------------+
| rcl                  |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| rhco3                |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| rso4                 |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| rno3                 |            | parameter needed to select the method                                           |
+----------------------+------------+---------------------------------------------------------------------------------+
| meth                 |            | method used like McNeal, Dunlap etc.                                            |
+----------------------+------------+---------------------------------------------------------------------------------+
| meth\_               |            | sub method used                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| rk20                 | µS/cm      | calculated conductivity at 20°C                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| ec25                 | mS/m       | calculated conductivity at 25°C                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| xecv                 | mS/m       | measured conductivity at 25°C                                                   |
+----------------------+------------+---------------------------------------------------------------------------------+
| ec25_xecv_sr         |            | standard residues of log transformed calculated and measured conductivity       |
+----------------------+------------+---------------------------------------------------------------------------------+
| pxecv                |            | transformation of xecv to calculate prinslabel                                  |
+----------------------+------------+---------------------------------------------------------------------------------+
| pec25                |            | transformation of ec25 to calculate prinslabel                                  |
+----------------------+------------+---------------------------------------------------------------------------------+
| prinslabel           |            | indicates whether the calculated conductivity does not match the measured value |
+----------------------+------------+---------------------------------------------------------------------------------+
| percentage_xecv_ec25 |            | percentage difference between measured and calculated conductivity              |
+----------------------+------------+---------------------------------------------------------------------------------+
| skat_san_sr          |            | standard residues of log transformed sum of kations or anions                   |
+----------------------+------------+---------------------------------------------------------------------------------+
| max_anion            | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| max_anion_name       |            | anion with the highest milliequivalents per liter                               |
+----------------------+------------+---------------------------------------------------------------------------------+
| max_kation           | meq/L      |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| max_kation_name      |            | kation with the highest milliequivalents per liter                              |
+----------------------+------------+---------------------------------------------------------------------------------+
| suspect              |            | none or max_anion or max_kation                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+
| percentage           |            | difference between script Stuyfzand and this script in %                        |
+----------------------+------------+---------------------------------------------------------------------------------+
|                      |            |                                                                                 |
+----------------------+------------+---------------------------------------------------------------------------------+

: Output headings

The data generally have a logarithmically transformed normal distribution. The higher values show higher differences but the relative differences remain constant at higher values. For example duplicate measurements may have a 10% error over a wide range of concentrations. The value of standard residues of log transformed data shows whether the data really deviate. 95% of the data fall between -2 and +2. When ec25_xecv_sr \> 2 the calculated conductivity is much higher than the measured one. When the skat_san_sr \< (-2) there is an excess of anions. When the calculated conductivity is much higher than the measured one and the ion balance has an excess of anions, then the anion with the maximum concentration of milliequivalents is suspect.

## Output style Stuyfzand

The output style Stuyfzand is a specialized output style to compare the results of the calculated conductivity k20 from the report of Stuyfzand with the calculated conductivity rk20 in this groundwaterconductivity::calculate_conductivity.R script. Both are expressed in micro Siemens/centimeter at 20°C. The percentage column is calculated as 100\*(rk20-k20)/k20 and rounded to zero digits. In the file Stuyfzand_minimal_output.csv the comparison between the report of Stuyfzand and our script is shown. Generally the difference is very small except for row 3, row 5 and row 7 all using the first method of Rossum and having very low conductivities . The five other rows using this method gave much more accurate results.

## Output style minimal

agafgh

## Add bicarbonate

aaaa

## Add phosphate

```{r}
data(veld)
data(put)
data(metingen)
data(filter)
```

Overzicht tabel 1 met veldgegevens

```{r}
#str(veld)
#head(veld)
```

Overzicht tabel 2 met put informatie

```{r}
#str(put)
#head(put)
```

Overzicht tabel 3 met putfilter informatie

```{r}
#str(filter)
#head(filter)
```

## Installatie KRWQCprotocol R-pakket

Het KRWQCprotocol wordt beschikbaar gesteld als R pakket. Gebruikers kunnen de functies uit dit pakket gebruiken in hun eigen scripts. Om het pakket te installeren moet vanuit R het volgende commando uitgevoerd worden:

```{r eval=FALSE}
#remotes::install_github("jspijker/KRWQCprotocol")
```

## Werking KRWQCprotocol R-pakket

De verschillende QC stappen uit het Protocol kunnen eenvoudig uitgevoerd worden door het aanroepen van de relevante functie (QC0a-QC5). Als invoer van de functie worden de relevante datatabellen meegegeven benodigd voor de testen. In de functie kan met het verbose argument opgegeven worden of de print van tussentijdse output gewenst is door deze op `TRUE` te zetten. In een aantal gevallen dienen extra invoervariabelen opgegeven te worden. Zie hiervoor de documentatie bij de functies, bijvoorbeeld `?KRWQCprotocol::QC0a`.

In de verschillende QC functies worden de testen uitgevoerd beschreven in het QC Protocol. Als bij deze stappen een afwijking geconstateerd wordt dan krijgt het monster, de parameter of de waarde een concept oordeel toegekend. Deze afwijkingen worden aan het eind van de functie als attribuut meegegeven aan het metingen bestand (`d_metingen`) en kunnen dus na het runnen van elke test bekeken worden.

*BELANGRIJK:* alle resultaten en concept oordelen worden per QC-test als attribuut aan het metingenbestand gehangen. Het is daarom noodzakelijk om voor elke test hetzelfde bestand als invoer te gebruiken, zodat alle test resultaten uiteindelijk als attribuut beschikbaar zijn onder eenzelfde bestand.

## Toepassing functies

Hieronder zullen de QC stappen 0-5 kort doorlopen worden om te laten zien hoe de functies toegepast worden en of aanvullende acties benodigd zijn.

Bij een viertal QC stappen moet de validatie door de grondwaterbeheerder zelf worden uitgevoerd. Het gaat om de stappen QC0c, QC0d, QC2b en QC2c. Om deze stappen in het R-pakket op te kunnen nemen moet voor deze functies eerst een invoerbestand gegenereerd worden met `QC0c_create_file`, idem voor de overige 3 functies. Na het invullen van deze bestanden dient het bestand onder dezelfde naam opgeslagen te worden zodat deze meegenomen kan worden in de functies QC0c, QC0d, QC2b en QC2c uit dit pakket. De resultaten worden net als bij de andere functies vervolgens als attribute toegekend aan het metingen bestand.

### QC0 - Meetopstelling

De stappen en functies onder QC0 worden op reguliere wijze uitgevoerd door het aanroepen van de betreffende functie en benodigde datasets, met uitzondering van QC0c en QC0d zie bovenstaande.
